Sensor Clusterization in D-optimal Design in Infinite Dimensional Bayesian Linear Inverse Problems

### 2023-01-03 latex/paper.tex
With corollary on pushforward/pullback and without a proper statement on k

\documentclass{amsart}

\input{definitions}

@@

@@

\begin{corollary}\label{cor:true target}
  Let $\obs$ with $m$ observations satisfy the nonlinear eigenvalue
  problem \eqref{eq:eigenproblem}. Let $\{\eta_i\}_{i=1}^{\infty}$
  eigenvalues of $\obs^*\obs$ and $\{\lambda_i\}_{i=1}^{\infty}$ the
  corresponding eigenvalues of $\fwd \prcov \fwd^*$. Then $k:=\rank
  \obs^*\obs \leq m$ and $\obs^*\obs$ has exactly $k$ positive
  eigenvalues.

  Without loss of generality, let $\eta_i > 0$ for $i\leq k$ and
  $\eta_i = 0$ for $i > k$. Then:
  \begin{equation*}
    \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i).
  \end{equation*}
  where $C$ dependes only on $\sigma^2$ and on
  $\{\lambda_i\}_{i=1}^{\infty}$.

\end{corollary}
\begin{proof}
  The first part is trivial. Since $k \leq m$, then $\obs^*\obs$ has
  at most $k$ nonzero eigenvalues $\eta_1, \dots, \eta_k$. Then:

  \begin{align}
    \begin{split}
      \tar(\obs) &= \frac12\log \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
      \obs \fwd \prcov^{1/2}\right )\\% \text{ (by definition)}\\
      %
      &= \frac12 \log \det \left (I + \sigma^{-2} \obs^* \obs \fwd
      \prcov\fwd^* \right ) \text{ (Sylvester's Determinant
      Theorem)}\\
      %
      %
      %
      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Corollary \ref{cor:same ev})} \\
      %
      %
      %
      %&=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )
      %
      %
      %
      &=\frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i).
      %
      %
      %% &=\frac12 \sum_{i=1}^{k} \log(\sigma^2\lambda_i^{-1} + \eta_i) - k\log \sigma + \frac12\sum_{i=1}^k \log \lambda_i.
    \end{split}
  \end{align}
\end{proof}

In light of Corollary \ref{cor:true target} we define:
\begin{definition}
  The \emph{utility proxy} $\bar{\tar}$ is defined for $\eta \in
  \mathbb{R}_+^k$, $\lambda \in \mathbb{R}_+^k$ and $\sigma^2 >0$ as:
  \begin{equation*}
    \bar{\tar}(\eta, \lambda, \sigma^2) := \sum_{i=1}^k \log(1 + \sigma^{-2}\lambda_i\eta_i).
  \end{equation*}
\end{definition}

\begin{observation}
  The utility proxy $\bar{\tar}$ is increasing in $\lambda$. Thus, in
  a D-optimal design $\lambda$ are the $k$ largest eigenvalues of
  $\fwd \prcov \fwd^*$.
\end{observation}

\begin{proposition}
  Fix $\lambda, \sigma^2$ and let
  $$
  \eta_i = \frac{1}{k} \left ( m - k \sigma^2 \lambda_i^{-1} + \sigma^2 \sum_{j=1}^k \lambda_j^{-1} \right ).
  $$

  If $\eta_i > 0,\ i=1,\dots, k$, then $\eta_1,\dots,\eta_k$ maximize
  the utility proxy $\bar{\tar}(\eta; \lambda, \sigma^2)$.
\end{proposition}
\begin{proof}
  Let $\Phi(\eta) = \sum \eta_i - m$. Using Lagrange multipliers:
  \begin{align*}
    \begin{split}
      \frac{\der}{\der \eta_i} \bar{\tar}(\eta, \lambda, \sigma^2) &=
      \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} \\
      %
      %
      %
      \frac{\der\Phi}{\der \eta_i} &= 1.
    \end{split}
  \end{align*}

  Denote $c := \sigma^{2}\lambda_i^{-1} + \eta_i$. Summing,
  substituting $\sum \eta_i = m$ and dividng by $k$:
  $$
  c = \frac{m}{k} + \sigma^2 \frac{1}{k} \sum_j \lambda_j^{-1}.
  $$

  Consequently:
  \begin{align*}
    \begin{split}
      \eta_i &= c - \sigma^2 \lambda_i^{-1} \\
      %
      %
      %
      &= \frac{m}{k} + \sigma^2 \frac{1}{k} \sum_j \lambda_j^{-1} - \sigma^2 \lambda_i^{-1} \\
      %
      %
      %
      &= \frac{1}{k} \left ( m - k \sigma^2\lambda_i^{-1} + \sigma^2 \sum_{j=1}^k \lambda_j^{-1} \right ).
    \end{split}
  \end{align*}
\end{proof}

%% We now turn our attention to maximizing $\bar{\tar}$ in $\eta$. The only
%% constraint on $\eta$ arises from the unit norm contraint on rows of
%% $\obs$, which implies that $\tr{\obs^*\obs} = \sum_i \eta_i= m$.
%% \begin{proposition}
%%   Fix $\lambda \in \mathbb{R}_+^k, \sigma^2 > 0$. Then
%%   $$
%%   \eta = \arg\max_{\sum_i \eta_i = m} \Phi(\eta; \lambda, \sigma^2)
%%   $$ 
%% \end{proposition}

%% \begin{proof}
%%   Lagrange multipliers:
%%   \begin{align*}
%%     \begin{split}
%%       \frac{d}{d\eta_i} \Phi(\eta; \lambda, \sigma^2) &= \frac{1}{\sigma^2\lambda_i^{-1} + \eta_i}, \\
%%       %
%%       %
%%       %
%%       \frac{d}{d\eta_i} (\sum_i \eta_i - m)  &= 1.
%%     \end{split}
%%   \end{align*}

%%   Denote the Lagrange multiplier $c$. Then
%%   \begin{align*}
%%     \begin{split}
%%       \sigma^2\lambda_i^{-1} + \eta_i &= \frac{1}{c}, \\
%%       %
%%       %
%%       %
%%       \sum_i \eta_i &= m.
%%     \end{split}
%%   \end{align*}

%%   and so

%%   \begin{align*}
%%     \begin{split}
%%       \sigma^2\sum_{i=1}^k\lambda_i^{-1} + m &= kc^{-1} \\
%%       %
%%     \end{split}
%%   \end{align*}

%% \end{proof}

%% Corollary \ref{cor:same ev} implies that instead of considering
%% $\obs$ when searching for a D-optimal design, we can restrict
%% ourselves to $k \leq m$ eigenvalues of $\obs^*\obs$ corresponding
%% to the first $k$ eigenvalues of $\fwd \prcov \fwd^*$.

The nonlinear eigenvalue problem \eqref{eq:eigenproblem} gives rise to
necessary conditions for D-optimality. Thus, in order to find the
global maximizer of $\tar$, we need to choose $k$, then choose $k$
eigenvalues $\lambda_1, \dots, \lambda_k$ and corresponding $\eta_1,
\dots, \eta_k$ to maximize $\sum_i \log( \sigma^2\lambda_i^{-1} +
\eta_i)$. It is not clear, however, if we can in fact choose
observation vectors to manifest any $\eta_1, \dots, \eta_k$
combination. The following lemma answers this in the
affirmitive. Moreover, the proof is constructive.

\begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M
  = m$, $m \geq k$. We can find $\func_j \in \R^k,j=1,\dots,m$
  with $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that
  $AA^t = M$.
\end{restatable*}

with the norm constraints on $\meas_j,j=1,\dots,m$. We would like to
make all $\eta_i$ as large as possible.  The problem of D-optimal
design has now been cast to finite dimensions. This justifies using
linear algebra. Lemma \ref{lemma:free} (a constructive proof is given
in the appendix) does exactly that.

Take any $\{\eta_i\}_{i=1}^{k}$, with $\sum_{i=1}^k \eta_i = m$. Let
$M = \diag(\eta_1,\dots,\eta_k)$. Lemma \ref{lemma:free} gives us $A$
with unit norm columns such that $AA^T = M$. Take $\meas_i
:=\sum_{j=1}^k A_{ji}\ev_j^*$. The constraints $\|\meas_i\| = 1,
i=1,\dots,m$ are satisified by construction. Verifying $\obs^*\obs
\ev_i = \eta_i \ev_i$ is not too difficult.

%% \begin{align*}
%%   \begin{split}
%%     \obs \ev_i &= (\meas_1,\dots,\meas_k)^t\ev_i\\
%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\ev_i\\
%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*\ev_i,\dots,\sum_{j=1}^kA_{jk}\ev_j^*\ev_i)^t\\
%%     &= (A_{i1},\dots,A_{ik})^t.
%%   \end{split}
%% \end{align*}

%% Now,

%% \begin{align*}
%%   \begin{split}
%%     \obs^*\obs \ev_i &= \obs^*(A_{i1},\dots,A_{ik})^t\\
%%     &=(A_{i1},\dots,A_{ik}) \obs \text{ (by \eqref{eq:obs*})}\\
%%     &=(A_{i1},\dots,A_{ik}) (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\\
%%     &=\sum_{l=1}^k A_{il} \sum_{j=1}^kA_{jl}\ev_j^* \\
%%     %&=\sum_{j=1}^k  \sum_{l=1}^kA_{il}A_{jl}\ev_j^* \\
%%     &=\sum_{j=1}^k  [AA^t]_{ij}\ev_j^* \\
%%     %&=\eta_i\ev_i^*\\
%%     &=\eta_i\ev_i^*. 
%%   \end{split}
%% \end{align*}

We have shown that we can construct $\obs$ such that it has
eigenvectors and eigenvalues $\{\ev_i, \eta_i\}_{i=1}^k$, as long as
$\sum_{i=1}^k \eta_i = m$. We now turn to understand which choice of
$\{\eta_i\}_{i=1}^k$ is optimal. By concavity of $\log$, the fastest
increase in the design criterion \eqref{eq:true target} is gained by
increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is smallest. So,
no weight should be given to (i.e. observation taken in direction of)
the $n$th eigenvector before $\sigma^2\lambda_i^{-1} + \eta_i \geq
\sigma^2\lambda_n^{-1},\ \forall i \neq n$. Thus, we should choose
$\{\eta_i\}_{i=1}^k$ as follows. First, order the eigenvalues
$\lambda_1 \geq \lambda_2 \geq \dots$. We should increase $\sigma^2
\lambda_1^{-1} + \eta_1$ until it equals
$\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} +
\eta_i,i=1,2$ are increased until $\sigma^2 \lambda_i^{-1} + \eta_i =
\sigma^2\lambda_3^{-1},i=1,2$ and so forth.

@@

    \end{axis}
  \end{tikzpicture}
  \caption{Characterizing a D-optimal design, compared to a
    sub-optimal design. Both designs are allowed three observations,
    which means the blue area has accumulated height of $3$ in both
    plots. A D-optimal design (left) increases precision (reduces
    uncertainty) where it is lowest. A sub-optimal design (right) does
    not. In this example, a observation is taken partly in the
    direction of the third eigenvector. The precision for this
    eigenvector is bigger even than the posterior precision of the
    D-optimal design on the left.}
  \label{fig:optimal vs not}
\end{figure}

@@

The argument above has the important consequence (illustrated in
figure \ref{fig:optimal vs not}) that the eigenvalues in the nonlinear
eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
perfect sense. We know the Lagrange multipliers $\xi_i,i=1,\dots,k$
tell us how much we gain by relaxing the constraints and there is no
reason for any eigenvector to give us more than any other (otherwise
we would put more weight on the better eigenvector, as can be
understood from the discussion in the following paragraph).

One can understand D-optimal designs as follows: Imagine each
eigenvector corresponds to a graduated lab cylinder (figure
\ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
liquid of $\sigma^2\lambda_i^{-1}$ volume units. This liquid
represnets the prior precision, weighted by observation noise. The
volume of the green liquid is increasing in $i$. A D-optimal design
with $m$ observations ignores cylinders indexed $m+1$ and above. It
has blue liquid of $m$ volume units at its disposal. It distributes
said liquid by repeatedly adding a drop to whatever cylinder currently
has the lowest level of liquid in it (lowest precision). The result of
such a procedure is illustrated in figure \ref{fig:optimal vs
  not}. The analysis conducted in this section is summarized in
Theorem \ref{thm:char} below.

\begin{theorem}[Characterizing D-optimal designs with vanishing model error]\label{thm:char}
  Let:
  \begin{enumerate}
  \item $\fwd, \prcov$ the forward and prior covariance operators, respectively.
  \item $m, \sigma^2$ the number of observations taken and observation
    noise variance, respectively.
  \item A D-optimal design operator $\obs$:
    $$
    \obs = \arg\max_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs)
    $$ 
  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
    in decreasing order of magnitude.
  %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
  \end{enumerate}
  Then:
  \begin{enumerate}
  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
  \item There exists $0 < k \leq m$ such that $\eta_i = 0$ for $i\geq
    k+1$ and $k\leq\rank \obs^*\obs$.
  \item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
  \item If $k < m$, then $\sigma^2\lambda_i^{-1} + \eta_i \leq
    \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
  \end{enumerate}
\end{theorem}

@@

\begin{lemma}[Increase due to a Observation]\label{lemma:design increase}
  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm := (\meas_1,\dots,\meas_{m-1})^t$. Then

>>>>

\documentclass{amsart}
\input{definitions}

@@

\begin{restatable}[D-optimal designs with vanishing model error]{theorem}{main}\label{thm:char}
  Let:
  \begin{enumerate}
  \item $\fwd:\hilp \to \hilo$ a linear forward operator,
    $\prcov:\hilp \to \hilp$ prior covariance operator, $\obs: \hilo
    \to \mathbb{R}^m$ observation operator, where $m \in \mathbb{N}$
    is the number of observations taken.
    \item $\sigma^2 \in \mathbb{R}_{+}$ observation noise variance,
      $\data = \obs \fwd \param + \eps$, where $\eps \in \mathbb{R}^m$
      is iid $\mathcal{N}(0, \sigma^2)$ noise, $\pr \sim
      \mathcal{N}(0, \prcov)$ prior measure, $\post$ the posterior
      measure.
  \item A D-optimality utility function
    \cite{AlexanderianGloorGhattas14}:
    \begin{align*}
      \begin{split}
        \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
        % 
        % 
        % 
        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
        \obs^* \obs \fwd \prcov^{1/2}).
    \end{split}
  \end{align*}
  \item A D-optimal design operator $\obs$:
    $$
    \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs)
    $$ 
  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
    in decreasing order of magnitude.
  %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
  \end{enumerate}

  Then:
  \begin{enumerate}
  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously
    diagonalizable.
  \item $k := \rank \obs^*\obs \leq m$.
  \item     
    \begin{equation*}
      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
    \end{equation*}
  \item 
    \begin{equation*}
      k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
      \lambda_j^{-1} \right \}.
    \end{equation*}
  \item
    \begin{equation*}
        \eta_i = \begin{cases}
          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
          0 & i > k 
        \end{cases}.
    \end{equation*}
  \end{enumerate}
\end{restatable}

@@

%% \begin{definition}
%%   Let $\{\lambda_i\}_{i=1}^{\infty}$ eigenvalues of $\fwd \prcov
%%   \fwd^*$ sorted in descending order. Let $\eta \in
%%   \mathbb{R}_+^m$. The \emph{utility proxy} $\bar{\tar}$ is defined
%%   as:
%%   \begin{equation*}
%%     \bar{\tar}(\eta) := \frac12 \sum_{i=1}^m \log(1 + \sigma^{-2}\lambda_i\eta_i).
%%   \end{equation*}
%% \end{definition}

\begin{corollary}\label{cor:true target}
  Let $\obs$ with $m$ observations satisfy the nonlinear eigenvalue
  problem \eqref{eq:eigenproblem}. Let $\{\eta_i\}_{i=1}^{\infty}$
  eigenvalues of $\obs^*\obs$ and $\{\lambda_i\}_{i=1}^{\infty}$ the
  corresponding eigenvalues of $\fwd \prcov \fwd^*$. Let $k:=\rank
  \obs^*\obs$. Without loss of generality, let $\eta_i > 0$ for $i\leq
  k$ and $\eta_i = 0$ for $i > k$. Then:
  \begin{enumerate}
    \item $k \leq m$ and $\obs^*\obs$ has exactly $k$ positive
      eigenvalues.
    \item
      \begin{equation*}
        \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i) = \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
      \end{equation*}
    \item Let $\obs$ a D-optimal design. Then $\eta_i > 0$ for
      eigenvectors corresponding to the $k$ largest $\lambda_i$.
  \end{enumerate}
\end{corollary}
\begin{proof}
  Part (1) is trivial. To see part (2) holds: 
  \begin{align}
    \begin{split}
      \tar(\obs) &= \frac12\log \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
      \obs \fwd \prcov^{1/2}\right )\\% \text{ (by definition)}\\
      %
      &= \frac12 \log \det \left (I + \sigma^{-2} \obs^* \obs \fwd
      \prcov\fwd^* \right ) \text{ (Sylvester's Determinant
      Theorem)}\\
      %
      %
      %
      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Corollary \ref{cor:same ev})} \\
      %
      %
      %
      %&=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )
      %
      %
      %
      &=\frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). 
      %
      %
      %% &=\frac12 \sum_{i=1}^{k} \log(\sigma^2\lambda_i^{-1} + \eta_i) - k\log \sigma + \frac12\sum_{i=1}^k \log \lambda_i.
    \end{split}
  \end{align}
  Part (3) holds since $\log$ is increasing and $\eta_i \geq 0$.
\end{proof}

\begin{proposition}\label{prop:optimizer}
  Let $\{\lambda_i\}_{i=1}^{\infty}$ the sorted eigenvectors of $\fwd
  \prcov\fwd^*$, $k\in \mathbb{N}^+$ and
  %% Let
  %%   \begin{equation}
  %%   k = \argmax \left \{ k:\lambda_k^{-1} \leq \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
  %%   \lambda_j^{-1} \right \}.
  %% \end{equation}

  \begin{equation}
  \eta_i = \begin{cases}
    \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & i \leq k \\
    0 & i > k 
  \end{cases}.
  \end{equation}

  Let $\obs$ an observation operator such that $\obs^*\obs$ has the
  same eigenvectors as $\fwd \prcov \fwd^*$. If $\eta_i > 0$ for
  $i\leq k$ and $\obs^*\obs$ has eigenvalues $\eta_i$, then $\obs$
  maximizes $\tar$ subject to $\sum_{i=1}^m \eta_i = m$ and $\rank
  \obs = k$.
\end{proposition}
\begin{proof}
  Below, we use a slight abuse of notation and cast the entire
  optimization problem of $\tar$ to the first $m$ eigenvectors of
  $\fwd\ \prcov \fwd^*$. We treat $\eta$ as if it is $m-$dimensional,
  since by \ref{}, $\eta_i = 0, i > m$. This abuse of notation gives
  us the freedom to think of the problem at hand in $m-$dimensions. We
  will now find the optimal $\eta_i$'s.

  First, let $A := \{i: \eta_i > 0\}$. A choice of $A$ where $\eta_i
  =0, \eta_j >0$ for some $1 \leq i < j \leq m$ is not optimal (just
  switch $\eta_j$ and $\eta_i$ and recall $\{\lambda_i\}$ are arranged
  in descending order). Thus, since $\rank \obs = k$ we conclude that,
  $\eta_i=0, i>k$.

  Let $\Phi(\eta) = \sum_{i=1}^k \eta_i - m$ and $\Omega_j(\eta) =
   -\eta_j$. Then
  \begin{align*}
    \begin{split}
      \frac{\partial \tar}{\partial \eta_i}  &=
       \frac12 \frac{\sigma^{-2}\lambda_i}{1 + \sigma^{-2} \lambda_i\eta_i} = \frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} \\
      %
      %
      %
      \frac{\partial\Phi}{\partial \eta_i} &= 1 \\
      %
      %
      %
      \frac{\partial \Omega_j}{\partial \eta_i} &= -\delta_{ij}      
    \end{split}
  \end{align*}

  From the KKT conditions, there are $\alpha, \beta_i$ such that for $i=1,\dots,m$:
  \begin{align}
    \begin{split}
      -\frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} + \alpha - \beta_i  &= 0 \\
      %
      %
      %
      \eta_i &\geq 0\\
      %
      %
      %
      \beta_i &\geq 0\\
      %
      %
      %
      \beta_i \eta_i &= 0\\
      %
      %
      %
      \sum_{i=1}^m \eta_i &= m 
    \end{split}
  \end{align}

  Then, for $1 \leq i \leq k$:
  \begin{align*}
    \begin{split}
      \beta_i &= 0\\
      %
      %
      %
      \sigma^{2}\lambda_i^{-1} + \eta_i  &= \frac{1}{2\alpha}%  \text{ for }  1 \leq i \leq k.\\
      %
      %
      %
      %% \sigma^{-2} \lambda_i &= 2\alpha - 2\beta_i, i \in B \\
    \end{split}
  \end{align*}

  Summing, substituting $\sum_{i=1}^m \eta_i = \sum_{i=1}^k \eta_i =
  m$, and dividng by $k$:
  $$
  \frac{1}{2\alpha} = \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j=1}^k \lambda_j^{-1}.
  %\Rightarrow \alpha = \left ( \frac{2m}{k} + \frac{2\sigma^2}{k}
  %\sum_{i\in A} \lambda_i^{-1} \right )^{-1}.
  $$

  Consequently:
  \begin{align} \label{eq:etas}
    \begin{split}
      \eta_i &= \frac{1}{2\alpha} - \sigma^2 \lambda_i^{-1}\\
      %
      %
      %
      &=\frac{m}{k} + \frac{\sigma^2}{k} \sum_{j=1}^k \lambda_j^{-1} - \sigma^2 \lambda_i^{-1}.%\\
      %
      %
      %
      %% &= \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j \in A, j\neq i} \lambda_j^{-1} - \frac{k-1}{k}\sigma^2 \lambda_i^{-1}\\
      %% ,\ i \in A.
    \end{split}
  \end{align}
\end{proof}
An immediate corollary from Proposition \ref{prop:} and the formula
for posterior eigenvalues is:
\begin{corollary}
  If $\obs$ is optimal and $k := \rank \obs$, then:
  \begin{enumerate}
    \item The pushforwad/pullback (??) $\fwd \post$ covariance is
      $((\fwd \prcov \fwd^*)^{-1} + \sigma^{\pm 2} \obs^*\obs)$.
    \item The eigenvalues of said operator are
      \begin{equation}
        \eta_i = \begin{cases}
          \frac{m}{k} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & i \leq k \\
          \lambda_i^{-1} & i > k 
        \end{cases}.
      \end{equation}
  \end{enumerate}
\end{corollary}
%% Thus, for a given $A$, we know which $\eta_i$ are optimal.

  %% Now, fix some $k$ and consider the constrained optimzation problem
  %% of maximizing $\tar$ subject to $\sum_{i=1}^k\eta_i=m$ (without the
  %% constraints $\eta_i \geq 0$). It is easy to verify via Lagrange
  %% multipliers that \eqref{eq:etas} are still optimal. Returning to our
  %% original constrained optimization problem, where $\eta_i \geq 0$,
  %% the optimal $\eta_i$'s are defined by \eqref{eq:etas}, where $k$
  %% takes the largest possible value for which \eqref{eq:etas} defines
  %% nonnegative $\eta_i$'s.

  %% By definition, $k$ satisfies $\eta_i > 0, i=1,\dots k$. Rearranging
  %% \eqref{eq:etas} for $\eta_i > 0$, we get that the following must
  %% hold:
  %% \begin{equation*} \label{eq:k}
  %% \lambda_i^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k}
  %% \sum_{j=1}^{k} \lambda_j^{-1}, i=1,\dots, k.
  %% \end{equation*}

  %% Since $\lambda_i$ are arranged in descending order, it suffices to
  %% verify that
  %% \begin{equation*}% \label{eq:k}
  %%   \lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
  %%   \lambda_j^{-1}.
  %% \end{equation*}

  %% Thus:
  %% \begin{equation*}
  %%   k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
  %%   \lambda_j^{-1} \right \}.
  %% \end{equation*}

We are almost ready to state and prove Theorem \ref{thm:char} that
characterizes D-optimal designs. We just need one final lemma, whose
proof is delegated to the appendix:
\begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M
  = m$, $m \geq k$. We can find $\func_j \in \R^k,j=1,\dots,m$
  with $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that
  $AA^t = M$.
\end{restatable*}

\main

\begin{proof}
  Part (1) was proved in Corollary \ref{cor:same ev}. Parts (2) and
  (3) were proved in Corollary \ref{cor:true target}. Parts (4) and
  (5) were proved in \ref{prop:optimizer}. The only missing part is
  showing that we can indeed find $\obs$ such that $\obs^*\obs$ has
  the required eigenvalues. To this end, we utilize Lemma
  \ref{lemma:free}: let $M = \diag(\eta_1, \dots, \eta_k)$, diagonal
  with respect to the first $k$ eigenvectors of $\fwd \prcov
  \fwd^*$. We take $\obs := A$ from the Lemma \ref{lemma:free}, and
  this concludes the proof.
\end{proof}
\begin{corollary}
  In a D-optimal design, the posterior precision for eigenvector $i$
  is $\eta_i + \sigma^2 \lambda_i^{-1} = \frac{m}{k} + \sigma^2
  \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1}$. The posterior precision is
  constant for $1 \leq i \leq k$.
\end{corollary}

@@

    \end{axis}
  \end{tikzpicture}
  \caption{As an illustration of Corollary \ref{cor:equal eigs}, a
    D-optimal design is compared with a sub-optimal design. Both
    designs are allowed three observations. Thus, the blue area has
    accumulated height of $3$ in both plots. A D-optimal design (left)
    increases precision (reduces uncertainty) where it is lowest. A
    sub-optimal design (right) does not. In this example, an
    observation is taken partly in the direction of the third
    eigenvector. The precision for this eigenvector is bigger even
    than the posterior precision of the D-optimal design on the left.}
  \label{fig:optimal vs not}
\end{figure}

@@

Here is a way to understand D-optimal designs with vanishing model
error: Imagine each eigenvector corresponds to a graduated lab
cylinder (Figure \ref{fig:cylinder}). Each cylinder $i$ is filled,
a-priori, with green liquid of $\sigma^2\lambda_i^{-1}$ volume
units. This liquid represnets the prior precision, weighted by
observation noise. The volume of the green liquid is increasing in
$i$. A D-optimal design with $m$ observations ignores cylinders
indexed $m+1$ and above. It has blue liquid of volume $m$ units at its
disposal. It distributes said liquid by repeatedly adding a drop to
whatever cylinder currently has the lowest level of liquid in it
(lowest precision). The result of such a procedure is illustrated in
Figure \ref{fig:optimal vs not}.

@@

\begin{lemma}[Increase due to an observation]\label{lemma:design increase}
  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
  (\meas_1,\dots,\meas_{m-1})^t$. Then
